The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-10-25 JStatSoft_forestinventory_master.rnw
formatting revision (2)

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable evidence-based information to assess the state and development of forests over time.  They typically consist of a random sample of plot locations in the forest that are assessed individually by hand. Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision. With respect to this objective, the application of multi-phase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient. While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata. This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in the \proglang{R} environment.}

@@

 % ... commential software tends to be expensive and lacks transparency making it less suitable for research purposes ...
 % ... free --> more suited for forest practice (limited financial ressources)

@@

\setlength{\parindent}{1em} % Einzug bei neuen Absätzen

@@

% execute everything that goes without echo:
<<echo=FALSE>>=
library(forestinventory)
@

>>>>

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable evidence-based information to assess the state and development of forests over time.  They typically consist of a random sample of plot locations in the forest that are assessed individually by hand. Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision. With respect to this objective, the application of multi-phase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient. While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multi-phase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata. This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in the \proglang{R} environment.}

@@

@@

\setlength{\parindent}{0em} % Einzug bei neuen Absätzen

@@

% execute everything that goes without echo:
<<echo=FALSE>>=
library("forestinventory")
@

### 2017-10-25 rnw/Analysis and Visualization.Rnw
formatting revision (2)

We often want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, which can be easily accomplished in \pkg{forestinventory} using the \code{estTable()} function. This function restructures the results from the \code{onephase()}, \code{twophase()} and \code{threephase()} objects and merges them into one single data set that provides the basis for further analysis. For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase extended pseudo synthetic and the pseudo synthetic estimators for the \code{grisons} data set:

>>>>

We often want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, which can be easily accomplished in \pkg{forestinventory} using the \code{estTable()} function. This function restructures the results from the \code{onephase()}, \code{twophase()} and \code{threephase()} objects and merges them into one single data set that provides the basis for further analysis. For demonstration purposes, we will first recalculate the one-phase estimator as well as the two-phase and three-phase EXTPSYNTH and PSYNTH estimator for the \code{grisons} data set:

### 2017-10-25 rnw/Calculation of Confidence Intervals.Rnw
formatting revision (2)

Converting the estimated variance into a 95\% confidence interval (CI) allows for a more practical interpretation of a point estimate's precision. The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value. In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sample design. Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples and calculate the 95\%  CI from each sample, then 95\% of the CIs are expected to contain the true population parameter. The confidence level $1-\alpha$ (e.g. 95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as coverage rate. The CI is also linked to hypothesis testing in that its associated point estimate is considered statistically different from any given value that lies outside the CI boundaries.

@@

where $\hat{Y}$ is the point estimate, $\hat{\var}(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model. In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots). In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function. For example, the 95\% confidence interval for the small area estimates by the extpsynth estimator (Section \ref{sec:twophase_sae}) are calculated by:

>>>>

Converting the estimated variance into a 95\% confidence interval (CI) allows for a more practical interpretation of a point estimate's precision. The correct interpretation of a CI is not that there is a 95\% probability that it contains the true value. In the design-based context, the true value of the population parameter we are trying to estimate, albeit unknown, is fixed and the sample is randomly generated under the sample design. Theoretically, if we were to repreatedly conduct the inventory using the same estimation method, estimator and auxiliary information under newly drawn random samples and calculate the 95\%  CI from each sample, then 95\% of the CIs are expected to contain the true population parameter. The confidence level $1-\alpha$ (e.g., 95\%) is thus the expected frequency or proportion of possible confidence intervals to contain the unknown population parameter under resampling and is therefore often also referred to as coverage rate. The CI is also linked to hypothesis testing in that its associated point estimate is considered statistically different from any given value that lies outside the CI boundaries.

@@

where $\hat{Y}$ is the point estimate, $\hat{\var}(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model. In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots). In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function. For example, the 95\% confidence interval for the small area estimates by the EXTPSYNTH estimator (Section \ref{sec:twophase_sae}) are calculated by:

### 2017-10-25 rnw/Introduction.Rnw
formatting revision (2)

% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.

@@

>>>>

% section: briefly give a review of existing R-packages for forest inventories (not much around)
While commercial enterprises have already been offering multi-phase inventory analyses, freely available software for application of multi-phase estimators has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} and the package \pkg{maSAE} by \citet{cullmann2016}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available. The motivation behind the \proglang{R} package \pkg{forestinventory} is to offer scientists and forest practitioners a comprehensive set of multi-phase estimation procedures that provide the advantages of a) full transparency of the implemented estimation methods, b) free access also for (small) forest enterprises who cannot afford commercial services, and c) implementation in an open source software that provides high flexibility to be incorporated in operational forest information systems.\par

The \pkg{forestinventory} package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management. The objectives of this article are to a) establish the link between the mathematical description of the estimators and their implementation in our package, b) illustrate their application to real-world inventory data sets and c) address special cases and demonstrate how the package-functions handle them.

% Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods to forestry has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} which provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available, which is the motivation behind the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

% Our thoughts:
 % ... commential software tends to be expensive and lacks transparency making it less suitable for research purposes ...
 % ... free --> more suited for forest practice (limited financial ressources)
 % ... not up to date: don't provide an extensive set of estimation methods (possibilities)
 % ... lacks the possibility to choose 
%  
% I'll try to keep it very general by emphasizing the pros of open-source software like ours
% that is: 1) full transparency of the methods, 2) free access also for (small) forest
% enterprises who maybe cannot afford commercial services, and 3) providing a huge set of estimators in a open source language that provides high flexibility to be incorporated in operational database system of forest services
% do you think that good?

@@

%% NOTE on commercial software:
%
% - SILVIATERRA: they call it "CruiseBoost" (https://silviaterra.com/bark/cruiseboost.html) --> means thinning the terrestrial inventory and
%                using remote sensing data instead , i.e. two-phase procedures. They say they are "model-assisted estimators". 
%                Costs "a low, per acre fee". 
%                Slogans "We guarantee that CruiseBoost will tighten up your confidence intervals or your money back". and 
%                "Your CruiseBoost estimate cannot be worse than the underlying traditional grid cruise" --> which is not true as we know
%
% - ForestMetrix: 
%
%
% Our thoughts:
 % ... commential software tends to be expensive and lacks transparency making it less suitable for research purposes ...
 % ... free --> more suited for forest practice (limited financial ressources)
 % ... not up to date: don't provide an extensive set of estimation methods (possibilities)
 % ... lacks the possibility to choose

### 2017-10-25 rnw/Threephase_Estimators and their Application.Rnw
formatting revision (2)

\begin{subequations}\label{eq:reg3p}
\begin{align}
\hat{Y}_{reg3p,EX}&=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
\nonumber \\&= (\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2} +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
\hat{Y}_{reg3p,NEX}&=\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
\nonumber \\&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2}  +
\hat{\bar{\pmb{Z}}}^t_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
\end{align}
\end{subequations}

@@

The synthetic (SYNTH) and pseudo synthetic estimator (PSYNTH) can be applied if no terrestrial samples are available in the small area, i.e., $n_{2,G}=0$. Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point estimate (Eq. \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (Eq. \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas. The point estimates are again potentially biased since $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$ for the full model residuals can not be ensured within small area $G$. Also the variance will be small but to the cost of ignoring the model uncertainties. Note that there is again no external variance formula for the synthetic and pseudo synthetic estimation.

\begin{subequations}\label{eq:3p_synth_psynth}
\begin{align}
\hat{Y}_{G,SYNTH,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2} \label{eq:pe_3p_synth} \\
\hat{Y}_{G,PSYNTH,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^t_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth} \\
\hat{\var}(\hat{Y}_{G,SYNTH,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_synth}\\
\hat{\var}(\hat{Y}_{G,PSYNTH,3p})& = \hat{\pmb{\alpha}}_2^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_psynth}
\end{align}
\end{subequations}

@@

The extended synthetic (EXTSYNTH) and extended pseudo synthetic estimator (EXTPSYNTH) ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e., $\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$. This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0. The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{\mathbb{Z}}^{(0)\top}(x),\pmb{Z}^{(1)\top}(x))$, where $\pmb{\mathbb{Z}}^{(0)\top}(x)=(\pmb{Z}^{(0)\top}(x), I_G^{\top}(x))$. In other words, when the extended option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for all observations in the input data frame (i.e., $s_0$). The regression coefficients, point estimates and variance estimates are calculated by replacing $\pmb{Z}$ with $\pmb{\mathbb{Z}}$ (and likewise $\pmb{Z}^{(0)}$ with $\pmb{\mathbb{Z}}^{(0)}$) into Eq. \ref{eq:normequ_3p}, \ref{eq:covar3p}, \ref{eq:pest_3p_small_psmall} and \ref{eq:3p_synth_psynth}. Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficients resulting from the random sample $s_2$.

@@

We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the extended synthetic and the synthetic estimator to the \code{grisons} data set. The setup is thus exactly the same as in the example for global three-phase estimation (Section \ref{sec:glob_est_3p}). However, this time will use the exact auxiliary mean of the mean canopy height variable (\code{mean}) and assume that we do not know the exact means of the remaining explanatory variables \code{stddev}, \code{max} and \code{q75}. We thus first define the true means for each small area just as we did in the \code{twophase()} example (Section \ref{sec:twophase_sae}):

@@

Three-phase small area estimation in the package can in general be applied by additionally specifying the \code{small_area} list argument. The exhaustive estimators can be called by optionally passing a \code{data.frame} containing the exact auxiliary means to the \code{exhaustive} argument. The extended synthetic estimator can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):

@@

The synthetic estimator can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.

@@

We see that the \code{threephase()}-function returns the sample sizes in the entire inventory area as well as within each small area. The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively. If we compare the two results, we see that the synthetic estimation again yields a much smaller variance than the extended synthetic estimation, but at the cost of a potential bias.

We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is non-exhaustively available but at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$. To do this, we additionally compute the extended pseudo synthetic estimates. As we can see, the exhaustive derivation of \code{mean} only yielded a slightly smaller variance.

>>>>

\begin{subequations}\label{eq:reg3p}
\begin{align}
\hat{Y}_{reg3p,EX}&=\frac{1}{\lambda(F)}\int_{F} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
\nonumber \\&= (\bar{\pmb{Z}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2} +
\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_ex} \\
\hat{Y}_{reg3p,NEX}&=\frac{1}{n_0}\sum_{x\in s_0} \pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2} + \frac{1}{n_1}\sum_{x\in s_1} (\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2}-\pmb{Z}^{(0)\top}(x)\hat{\pmb{\alpha}}_{s_2}) + \frac{1}{n_2}\sum_{x\in s_2}(Y(x)-\pmb{Z}^{\top}(x)\hat{\pmb{\beta}}_{s_2})
\nonumber \\&=(\hat{\bar{\pmb{Z}}}^{(0)}_0-\hat{\bar{\pmb{Z}}}^{(0)}_1)^{\top}\hat{\pmb{\alpha}}_{s_2}  +
\hat{\bar{\pmb{Z}}}^{\top}_1\hat{\pmb{\beta}}_{s_2} \label{eq:reg3p_nex}
\end{align}
\end{subequations}

@@

The synthetic (SYNTH) and pseudo synthetic (PSYNTH) estimator can be applied if no terrestrial samples are available in the small area, i.e., $n_{2,G}=0$. Consequently, the residual correction and the residual variation term of the full model can no longer be applied and drops from the point estimate (Eq. \ref{eq:pe_3p_synth} and \ref{eq:pe_3p_psynth}) and g-weight variance (Eq. \ref{eq:var_3p_reg_synth} and \ref{eq:var_3p_reg_psynth}) formulas. The point estimates are again potentially biased since $\frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$ for the full model residuals can not be ensured within small area $G$. Also the variance will be small but to the cost of ignoring the model uncertainties. Note that there is again no external variance formula for the synthetic and pseudo synthetic estimation.

\begin{subequations}\label{eq:3p_synth_psynth}
\begin{align}
\hat{Y}_{G,SYNTH,3p}&=(\bar{\pmb{Z}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2} \label{eq:pe_3p_synth} \\
\hat{Y}_{G,PSYNTH,3p}&=(\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}-\hat{\bar{\pmb{Z}}}^{(0)}_{1,G})^{\top}\hat{\pmb{\alpha}}_2 +
\hat{\bar{\pmb{Z}}}^{\top}_{1,G}\hat{\pmb{\beta}}_{s_2}\label{eq:pe_3p_psynth} \\
\hat{\var}(\hat{Y}_{G,SYNTH,3p})& =\frac{n_2}{n_1}\bar{\pmb{Z}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\bar{\pmb{Z}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_synth}\\
\hat{\var}(\hat{Y}_{G,PSYNTH,3p})& = \hat{\pmb{\alpha}}_2^{\top}\hat{\pmb{\Sigma}}_{\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}}\hat{\pmb{\alpha}}_2 +
\frac{n_2}{n_1}\hat{\bar{\pmb{Z}}}^{(0)\top}_{0,G}\hat{\pmb{\Sigma}}_{\hat{\pmb{\alpha}}_{s_2}}
\hat{\bar{\pmb{Z}}}^{(0)}_{0,G}+(1-\frac{n_2}{n_1})\hat{\bar{\pmb{Z}}}_{1,G}^{\top}\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}
\hat{\bar{\pmb{Z}}}_{1,G} \label{eq:var_3p_reg_psynth}
\end{align}
\end{subequations}

@@

The extended synthetic (EXTSYNTH) and extended pseudo synthetic (EXTPSYNTH) estimator ensures that the residuals of the full model over both the entire inventory area $F$ and the small area $G$ are zero at the same time, i.e., $\frac{1}{n_{2}}\sum_{x \in s_{2}}\hat{R}(x) = \frac{1}{n_{2,G}}\sum_{x \in s_{2,G}}\hat{R}(x)=0$. This is again realized by extending the vector of explanatory variables by a binary categorical indicator variable $I_G(x)$ which takes the value 1 if the observation lies inside the small area $G$ and is otherwise set to 0. The extended auxiliary vector is thus defined as $\pmb{\mathbb{Z}}^{\top}(x)=(\pmb{\mathbb{Z}}^{(0)\top}(x),\pmb{Z}^{(1)\top}(x))$, where $\pmb{\mathbb{Z}}^{(0)\top}(x)=(\pmb{Z}^{(0)\top}(x), I_G^{\top}(x))$. In other words, when the extended option is chosen, \pkg{forestinventory} automatically adds the binary indicator variable for the desired small area for all observations in the input data frame (i.e., $s_0$). The regression coefficients, point estimates and variance estimates are calculated by replacing $\pmb{Z}$ with $\pmb{\mathbb{Z}}$ (and likewise $\pmb{Z}^{(0)}$ with $\pmb{\mathbb{Z}}^{(0)}$) into Eq. \ref{eq:normequ_3p}, \ref{eq:covar3p}, \ref{eq:pest_3p_small_psmall} and \ref{eq:3p_synth_psynth}. Just as in the two-phase case, the resulting point estimates are now unbiased and have an associated g-weight variance that accounts for the variability of the regression coefficients resulting from the random sample $s_2$.

@@

We will demonstrate the use of three-phase small area estimation in the package \pkg{forestinventory} by applying the EXTSYNTH and SYNTH estimator to the \code{grisons} data set. The setup is thus exactly the same as in the example for global three-phase estimation (Section \ref{sec:glob_est_3p}). However, this time will use the exact auxiliary mean of the mean canopy height variable (\code{mean}) and assume that we do not know the exact means of the remaining explanatory variables \code{stddev}, \code{max} and \code{q75}. We thus first define the true means for each small area just as we did in the \code{twophase()} example (Section \ref{sec:twophase_sae}):

@@

Three-phase small area estimation in the package can in general be applied by additionally specifying the \code{small_area} list argument. The exhaustive estimators can be called by optionally passing a \code{data.frame} containing the exact auxiliary means to the \code{exhaustive} argument. The EXTSYNTH estimator can be applied by setting the argument \code{unbiased} to \code{TRUE} (default):

@@

The SYNTH estimator can be applied by changing the argument \code{unbiased} to \code{FALSE}, which causes the function to not apply a bias correction in the respective small area.

@@

We see that the \code{threephase()}-function returns the sample sizes in the entire inventory area as well as within each small area. The value \code{Inf} for \code{n0G} indicates that the explanatory variables at $s_0$ sample locations used in the reduced model were in our case derived exhaustively. If we compare the two results, we see that the SYNTH estimation again yields a much smaller variance than the EXTSYNTH estimation, but at the cost of a potential bias.

We can also analyse how the exhaustive derivation of \code{mean} performed compared to the case where \code{mean} is non-exhaustively available but at a very large $s_0$ phase with $n_{0,G}>>n_{1,G}$. To do this, we additionally compute the EXTPSYNTH estimates. As we can see, the exhaustive derivation of \code{mean} only yielded a slightly smaller variance.
